---
title: "disttools"
author: "Zachary Colburn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Functions that determine the distances between objects typically return either an object of class 'dist' or a matrix. When a matrix is returned, there is wasted memory because a distance matrix is symmetric across the diagonal (top left to bottom right). On the other hand, a 'dist' object, while more memory efficient, is inconvenient because its distance information can only be accessed using a one dimensional index. Accessing the distance between any two data points therefore requires either coercing the 'dist' object to a matrix using *as.matrix*. or calculating the one dimensional index that corresponds to the two data points of interest. When working with very large 'dist' objects the former may not be feasible. The package *disttools* provides a simple interface for accessing the distances contained within a distance matrix without the memory overhead of *as.matrix* coercion.


The function *get_dists* allows distances to be accessed for different pairs of points in two ways, demonstrated below.
```{r}
# Load the package.
library(disttools)

# Create some data to play with.
mat <- matrix(rnorm(10), ncol = 2)
mat_dists <- dist(mat)
indices <- matrix(c(1,2,3,2,4,4), ncol = 2, byrow = TRUE)

# Retrieve distances using the matrix-based method.
get_dists(mat_dists, indices)
```


Alternatively, the data can be accessed using a pair of vectors.
```{r}
# Create vectors i and j from the above data.
i <- indices[,1]
j <- indices[,2]

# Retrieve distances using the paired vectors method.
get_dists(mat_dists, i, j)
```